О множествах точек на плоскости с целочисленными расстояниями

### 2019-05-09 pps-linear-bound/pps-linear-lower-bound.tex
Amsthm completed

\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]

@@

\paragraph{Construction 1.}
Let us choose a positive integer $k > 1$ and take
\begin{equation}
	a = 2^{2^k} - 1
	.
\end{equation}
Then
\begin{equation}
	\label{eq:a_equiv_3_mod_4}
	a \equiv 3 \mod 4
\end{equation}
and, moreover,
\begin{multline}
	a = \left(2^{2^{k-1}}\right)^2 - 1
	=
	\left(2^{2^{k-1}} + 1\right) \left(2^{2^{k-1}} - 1\right)
	=
	\\=
	\left(2^{2^{k-1}} + 1\right) \left(2^{2^{k-2}} + 1\right) \cdot ... \cdot \left(2^{2^1} + 1\right) \left(2^2 - 1\right)
	.
\end{multline}
We take $d_j = 2^{2^j} + 1$ for $1 \leq j \leq k-1$.
Then $d_j \equiv 1 \mod 4$.

@@

So, $M = \{ M_J, N, H_{1,2}\}$ is indeed a planar integral point set of $2^{k-1}+3$ points,
and distance 1 occurs in $M$.
Reminding that $k$ can be taken arbitrary large, we conclude that we can construct
a planar integral point set of arbitary large count of points so that the distance 1 occurs in that set.

@@

\paragraph{Theorem 3.1.}
For every $n\geq 3$ there is a planar intergal set $M$ of $n$ points
such that for some $M_1,M_2 \in M$ equality $|M_1 - M_2|=1$ holds.
This set consists of $n-1$ points, including $M_1$ and $M_2$, on a straight line and one point out of the line.

And vice versa, if $M$ is a planar integral point set of $n$ points
such that for some $M_1,M_2 \in M$ equality $|M_1 - M_2|=1$ holds,
then $M$ consists of $n-1$ points, including $M_1$ and $M_2$, on a straight line and one point out of the line.
Moreover, if $n > 3$, then the distance 1 occurs in $M$ only once.

Using the ``blowing up'' procedure described in~\cite[theorem 1.3]{kurz2008bounds},
we can construct an integral point set $M$ in $m$-dimensional Euclidean space, $m\geq 3$,
with distance 1 occuring in it.
If we have an integral point set with distance 1 occuring,
then we can easily stretch it to turn 1 into the desired distance.
These facts, together with theorem 3.1, give the following

@@

\paragraph{Definition.}
We will call an integral point set $M$ \textit{irreducible}, if the greatest common divisor
of all the distances occuring in $M$ is 1.

@@

\paragraph{Hypothesis 3.1.}
For every $m \geq 2$, $n \geq m+1$, $d \geq 1$ there exists an \textbf{irreducible}
integral point set $M$ of $n$ points in $m$-dimensional Euclidean space
where $M$ contains points $M_1$ and $M_2$ such that distance between $M_1$ and $M_2$ is exactly $d$.

For $m \geq 3$, hypothesis 3.1 can be easily proved:
indeed, we can just take a facher integral point set according to construction 1
with sufficiently large height and then apply the ``blowing up'' procedure described in~\cite[Theorem 1.3]{kurz2008bounds},
replacing the out-of-line point with an $(m-2)$-dimensional simplex with proper side length.
That simplex will consist of $m-1$ points;
we can throw out up to all the points $M_J$ and still have points $H_{1,2}$ with distance 1.
So, the set consists of $m+1$ or more points and has distance 1 occuring in it.
As distance 1 occurs in the obtained set, it is irreducible.

So, hypothesis 3.1 needs proof only for $m=2$.

>>>>

\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{hypothesis}[theorem]{Hypothesis}
\newtheorem{lemma}[theorem]{Lemma}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{construction}[theorem]{Construction}

@@

\begin{construction}
	\label{con:planar_set_with_minimeter_1}
	Let us choose a positive integer $k > 1$ and take
	\begin{equation}
		a = 2^{2^k} - 1
		.
	\end{equation}
	Then
	\begin{equation}
		\label{eq:a_equiv_3_mod_4}
		a \equiv 3 \mod 4
	\end{equation}
	and, moreover,
	\begin{multline}
		a = \left(2^{2^{k-1}}\right)^2 - 1
		=
		\left(2^{2^{k-1}} + 1\right) \left(2^{2^{k-1}} - 1\right)
		=
		\\=
		\left(2^{2^{k-1}} + 1\right) \left(2^{2^{k-2}} + 1\right) \cdot ... \cdot \left(2^{2^1} + 1\right) \left(2^2 - 1\right)
		.
	\end{multline}
	We take $d_j = 2^{2^j} + 1$ for $1 \leq j \leq k-1$.
	Then $d_j \equiv 1 \mod 4$.

@@

	So, $M = \{ M_J, N, H_{1,2}\}$ is indeed a planar integral point set of $2^{k-1}+3$ points,
	and distance 1 occurs in $M$.
	Reminding that $k$ can be taken arbitrary large, we conclude that we can construct
	a planar integral point set of arbitary large count of points so that the distance 1 occurs in that set.
\end{construction}

@@

\begin{theorem}
	\label{thm:minimeter_1_planar}
	For every $n\geq 3$ there is a planar intergal set $M$ of $n$ points
	such that for some $M_1,M_2 \in M$ equality $|M_1 - M_2|=1$ holds.
	This set consists of $n-1$ points, including $M_1$ and $M_2$, on a straight line and one point out of the line.

	And vice versa, if $M$ is a planar integral point set of $n$ points
	such that for some $M_1,M_2 \in M$ equality $|M_1 - M_2|=1$ holds,
	then $M$ consists of $n-1$ points, including $M_1$ and $M_2$, on a straight line and one point out of the line.
	Moreover, if $n > 3$, then the distance 1 occurs in $M$ only once.
\end{theorem}

Using the ``blowing up'' procedure described in~\cite[theorem 1.3]{kurz2008bounds},
we can construct an integral point set $M$ in $m$-dimensional Euclidean space, $m\geq 3$,
with distance 1 occuring in it.
If we have an integral point set with distance 1 occuring,
then we can easily stretch it to turn 1 into the desired distance.
These facts, together with Theorem~\ref{thm:minimeter_1_planar}, give the following

@@

\begin{definition}
	We will call an integral point set $M$ \textit{irreducible}, if the greatest common divisor
	of all the distances occuring in $M$ is 1.
\end{definition}

@@

\begin{hypothesis}
	\label{hyp:irreducible_planar}
	For every $m \geq 2$, $n \geq m+1$, $d \geq 1$ there exists an \textbf{irreducible}
	integral point set $M$ of $n$ points in $m$-dimensional Euclidean space
	where $M$ contains points $M_1$ and $M_2$ such that distance between $M_1$ and $M_2$ is exactly $d$.
\end{hypothesis}

For $m \geq 3$, Hypothesis~\ref{hyp:irreducible_planar} can be easily proved:
indeed, we can just take a facher integral point set according to Construction~\ref{con:planar_set_with_minimeter_1}
with sufficiently large height and then apply the ``blowing up'' procedure described in~\cite[Theorem 1.3]{kurz2008bounds},
replacing the out-of-line point with an $(m-2)$-dimensional simplex with proper side length.
That simplex will consist of $m-1$ points;
we can throw out up to all the points $M_J$ and still have points $H_{1,2}$ with distance 1.
So, the set consists of $m+1$ or more points and has distance 1 occuring in it.
As distance 1 occurs in the obtained set, it is irreducible.

So, Hypothesis~\ref{hyp:irreducible_planar} needs proof only for $m=2$.
